---
title: "SDS 315 - HW3"
author: "Jason George (jg82397)"
date: "https://github.com/JasonGeorge978/SDS-315"
output:
  pdf_document: default
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=10, warning=FALSE, message=FALSE, tidy=TRUE, echo=FALSE, tidy.opts=list(width.cutoff=60))
```

```{r}
library(tinytex)
library(tidyverse)
library(ggplot2)
library(mosaic)
nbc <- read_csv("nbc_pilotsurvey.csv")
gasprices <- read_csv("gasprices.csv")
sclass <- read_csv("sclass.csv")
ebay <- read_csv("ebay.csv")
```

\newpage

# **Problem. 1**

### **Part A**

```{r}
ggplot(gasprices) + geom_boxplot(aes(x=Competitors, y= Price),fill="tomato") + theme_minimal() + labs(title="Price of Gas vs Competitors in Area",x= "Competitors", y= "Price($)")
gasprices$Competitors_bool <- (gasprices$Competitors == "Y")
boot_competitors = do(10000)* diffmean(Price ~ Competitors_bool, data=resample(gasprices))
confint(boot_competitors, level = 0.95)
```

Claim:
Gas stations charge more if they lack direct competition in sight.

Evidence: The boxplot shows no competitor in sight has a higher median and range. The difference of prices with or without competitors ranges from -0.055 to 0.007 with a 95% confidence interval. 

Conclusion:
The theory is not supported by the data as the interval crosses 0, suggesting no strong evidence of a statistical significance between the two much less an increase in cost.

\newpage

### **Part B**

```{r}
ggplot(gasprices) + geom_point(aes(x=Income,y=Price),fill="tomato") + theme_minimal() + labs(title="Price of Gas vs Income of Area", x="Income", y="Price($)")
#summary(gasprices$Income)
#summary tells us 3rd Quartile is 70k
gasprices <- mutate(gasprices,is_high= Income > 70000)
ggplot(gasprices) + geom_boxplot(aes(x=is_high, y= Price,fill=is_high)) + theme_minimal() + labs(title="Price of Gas vs High and Low Income", x="High or Low Income", y="Price($)",fill="High Income") + theme_minimal()
boot_income = do(10000)* diffmean(Price ~ is_high , data=resample(gasprices)) 
confint(boot_income, level = 0.95)
```

Claim:
The richer the area, the higher the gas prices

Evidence: 
The boxplot shows that areas classified as High Income have higher gas prices and the scatter plot shows majority of high price gas occuring at higher incomes. The 95% confidence interval for difference in price was from 0.013 to 0.085.

Conclusion:
The data supports the claim that prices are a little bit higher in higher income areas as the confidence interval lies above 0. 
\newpage

### **Part C**

```{r}
ggplot(gasprices) + geom_boxplot(aes(x=Stoplight, y= Price),fill="tomato") + theme_minimal() + labs(title="Price of Gas vs Pressence of Stoplight", x="Pressence of Stoplight", y="Price($)")
gasprices$Stoplight_bool <- gasprices$Stoplight == "Y"
boot_stoplight = do(10000)* diffmean(Price ~ Stoplight_bool , data=resample(gasprices))
confint(boot_stoplight, level = 0.95)
```

Claim:
Gas stations at stoplights charge more

Evidence:
The boxplot shows overlapping price distributions for gas stations based on stoplight pressence. The 95% confidence interval for the difference in mean price ranges from -0.0384 to 0.0302.

Conclusion:
The data does not support the claim that gas stations are stoplights charge more as the interval includes 0, meaning no strong evidence of a correlation. 

\newpage

### **Part D**

```{r}
ggplot(gasprices) + geom_boxplot(aes(x=Highway, y= Price),fill="tomato") + theme_minimal() + labs(title="Price of Gas vs Pressence of Highway Nearby", x="Pressence of Highway", y="Price($)")
gasprices$Highway_bool <- gasprices$Highway == "Y"
boot_highway = do(10000)* diffmean(Price ~ Highway_bool , data=resample(gasprices))
confint(boot_highway, level = 0.95)
```

Claim:
Gas stations with direct highway access charge more

Evidence: 
The boxplot shows that gas stations near a highway have higher prices than without. The 95% confidence interval for the difference ranges from 0.009 to 0.081.

Conclusion:
Since the interval lies above 0, the data supports the claim that gas stations with highway access charge more. 

\newpage

### **Part E**

```{r}
gasprices$is_shell <- gasprices$Brand == "Shell"
ggplot(gasprices) + geom_boxplot(aes(x=is_shell, y= Price),fill="tomato") + theme_minimal() + labs(title="Price of Gas vs Pressence of Stoplight", x="Pressence of Stoplight", y="Price($)")
boot_shell = do(10000)* diffmean(Price ~ is_shell , data=resample(gasprices))
confint(boot_shell, level = 0.95)
```

Claim:
Shell charges more than all other non-Shell brands.

Evidence: 
The boxplot shows overlapping price distribution. The estimated difference was 0.027, with a 95% confidence interval ranging from -0.009 to 0.065. 

Conclusion:
The data does not support the claim that Shell charges more as the interval crosses 0, suggesting the difference could be 0. 
\newpage

# **Problem. 2**


### **Part A**

```{r}
part_A <- filter(sclass,year == 2011 & trim == "63 AMG")
boot_2011_AMG = do(10000)* mean(~ mileage, data=resample(part_A))
confint(boot_2011_AMG, level = 0.95)
```

The average mileage of 2011 S-Class 63 AMGs that hit the used-car markert with a 95% confidence interval is between 26226 and 31790.

### **Part B**

```{r}
part_B <- filter(sclass,year == 2014 & trim == "550")

part_B <- mutate(part_B, isBlack = color=="Black")

boot_2014_550 = do(10000)* prop(~ isBlack, data=resample(part_B))

confint(boot_2014_550, level = 0.95)

```

The proportion of all 2014 S-Class 550s that were painted Black with a 95% confidence interval is between 0.417 and 0.453

\newpage

# **Problem. 3**


### **Part A**


```{r}
P3_both <- filter(nbc, Show == "Living with Ed" | Show == "My Name is Earl")

boot_both <- do(10000) * diffmean(Q1_Happy ~ Show, data = resample(P3_both))
confint(boot_both, level = 0.95)

```

Question:
Is there evidence that one show consistently produces
a higher mean Q1_Happy response among viewers?

Approach: 
First I filtered the NBC dataset to only include the shows "Living with Ed" and "My Name is Earl." Then I ran created a bootstrap, running 10000 simulations to then find the confidence interval for the difference. 

Results:
With a 95% confidence interval the difference of mean Q1_happy scores ranges from -0.3984323 to 0.09894481.

Conclusion:
The claim is not supported by data, as the interval passes through 0, we cant find a difference, meaning viewers might find both equally as funny. 

### **Part B**


```{r}
P3_both2 <- filter(nbc, Show == 'The Biggest Loser' | Show == 'The Apprentice: Los Angeles')

boot_both2 <- do(10000) * diffmean(Q1_Annoyed ~ Show, data = resample(P3_both2))
confint(boot_both2, level = 0.95)
```

Question: 
Is there evidence that one show consistently produces a higher mean Q1_Annoyed response among viewers?

Approach:
First, the dataset was filtered to include only “The Biggest Loser” and “The Apprentice: Los Angeles.” A bootstrap with 10,000 simulations was performed to generate a 95% confidence interval for the difference in mean Q1_Annoyed ratings between these two shows.

Results:
The 95% confidence interval for the difference in mean Q1_Annoyed ratings ranges from -0.5213588 to -0.01315503, with an estimate of -0.270997.

Conclusion:
Because the confidence interval is below 0, it indicates "The Biggest Loser" had a lower mean, meaning "The Apprentice: Los Angeles" is more annoying on average. 

### **Part C**

```{r}
P3_both3 <- filter(nbc, Show == 'Dancing with the Stars')

P3_both3 <- mutate(P3_both3, Above_4 = Q2_Confusing >= 4 )
boot_both3 <- do(10000) * prop(~Above_4, data = resample(P3_both3))
confint(boot_both3, level = 0.95)

```

Question:
What proportion of American TV watchers would we expect to give a
response of 4 or greater to the “Q2_Confusing” question?

Approach:
First the dataset was filtered to include only responses for "Dancing with the Stars." Then a new variable called Above_4, to indicate whether the confusing score was equal to or higher than 4. Then a boostrap with 10000 simulations find the confidence interval. 

Results:
The estimated proportion of viewers who rated the show as confusing is 0.0773. The 95% confidence interval ranged from 0.0387 to 0.1160. 

Conclusion:
While some viewers do find "Dancing with the Stars" confusing, the data shows the proportion of those who do is relatively small. 

\newpage

# **Problem. 4**

```{r}
ebay <- mutate(ebay, rev_ratio = rev_after / rev_before)

boot_ebay <- do(10000) * diffmean(rev_ratio ~ adwords_pause, data = resample(ebay))
confint(boot_ebay, level = 0.95)
```

Question:
Whether the revenue
ratio is the same in the treatment and control group?

Approach:
First I added a new column named rev_ratio which would display the revenue ration per. Then I made 10000 Monte Carlo simulations to then find the confidence interval of the difference of rev_ratio within treatment or control group. 

Results:
The estimated difference between the two is -0.05 with a 95% confidence interval from -0.09061933 to -0.01323848.

Conclusion:
The confidence values range not including 0 indicates evidence they have a difference. The negative estimate suggests pausing AdWords would lead to lower revenue ratios. 
