---
title: "SDS 315 - HW 7"
author: "Jason George (jg82397)"
date: "https://github.com/JasonGeorge978/SDS-315"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=5, fig.width=10, warning=FALSE, message=FALSE, tidy=TRUE, echo=FALSE, tidy.opts=list(width.cutoff=60))
```

```{r}
library(tinytex)
library(mosaic)
library(ggplot2)
library(MatchIt)

armfold = read.csv('armfold.csv')
```

\newpage

## **Problem 1: Armfolding**


### **Part A**

```{r}
#table(armfold$Sex)


males <- armfold[armfold$Sex == "Male", ]

males_prop = sum(males$LonR_fold == 1) / nrow(males)

females <- armfold[armfold$Sex == "Female", ]

females_prop = sum(females$LonR_fold == 1) / nrow(females)

```
The number of males in the dataset are 106 and the number of females is 111.

The sample proportion of males who folded their left arm on top is 0.47.

The sample proportion of females who folded their left arm on top is 0.42.

### **Part B**

```{r}
prop_diff = males_prop - females_prop
```

The observed difference between males and females is 0.048.

### **Part C**


```{r}
# Calculating standard error
standard_error <- sqrt((males_prop * (1 - males_prop )) / nrow(males) + (females_prop * (1 - females_prop)) / nrow(females))
```

The formula for standard error for the difference in proportions is sqrt((p1(1-p1))/n1 + (p2(1-p2))/n2).

The values I plugged into the formula are p1(proportion of males who folded their left arm on top), n1(total males), p2(proportion of females who folded their left arm on top), and n2(total females)



```{r}
# Assign z*
z_star = 1.96
```
A z* value of 1.96 was chosen because we are looking for the 95% confidence interval


```{r}
# Confidence Interval
lower = prop_diff - z_star * standard_error
upper = prop_diff + z_star * standard_error
```



```{r}
# R base function
male_sum <- sum(males$LonR_fold == 1)
female_sum <- sum(females$LonR_fold == 1)

prop = prop.test(x=c(male_sum, female_sum),n=c(nrow(males),nrow(females)),conf.level = 0.95)

#print(prop)
```
Our calculated interval was -0.08 to 0.18

The R in-built function gave us an interval of -0.09 to 0.19

### **Part D**

If we were to repeat this sampling process infinitely, then we would expect that 95% of the difference in proportion between males and females will be from -0.08 to 0.18

### **Part E**

The standard error value of 0.067 we calculated gives us the estimated variability in the difference of proportions between makes and females putting their left hand over their right hand due to random sampling and chance.

### **Part F**

The term sampling distribution describes the distribution of differences that we would see if we repeatedly took random samples of males and females from the population and calculated the diff proportion. What is varying from sample to sample is the difference porportions of males and females fold their arms with the left arm on top. The thing that stays the same is the population proportions and and sampling process. 

### **Part G**

The mathematical result that justifies using a normal distribution to approximate the sampling distribution is the Central Limit Theorem. It describes that if a large number of random samples from any population is taken the sampling distribiton will become a normal distribution if the sample size is large enough. 

### **Part H**

Despite the interval including 0. We should not rule out the possibility that there is no difference between sex in arm folding. Also with such a wide difference from -1% to 30% there seems to be be a possibility of a large difference with left arm on top for men. This interval should not conclusively prove that there is not a difference instead may suggest there is. 

### **Part I**

Yes the confidence interval would vary across different samples. Due to the natural random differences in selections and sampling the interval would include slightly difference proportions.If we had all of these intervals we could assume 95% of those intervals contain the actual proportion difference.


\newpage

## **Problem 2: Get out the vote**

```{r}
turnout = read.csv('turnout.csv')
```

### **Part A**

```{r}
gotv_1998 <- turnout[turnout$GOTV_call == 1,]

non_gotv <- turnout[turnout$GOTV_call == 0, ]

gotv <- mean(gotv_1998$voted1998)
non <- mean(non_gotv$voted1998)

gotv_total <- nrow(gotv_1998)
non_total <- nrow(non_gotv)

diff_prop <- gotv - non

se_gotv <- sqrt((gotv * (1 - gotv)) / gotv_total + (non * (1 - non)) / non_total)

turn_lower <- diff_prop - z_star * se_gotv
turn_upper <- diff_prop + z_star * se_gotv



```

The proportion of those receiving a GOTV call who voted in 1998 was 0.65.

The sample proportion of those not receiving a GOTV call who voted in 1998 was 0.44.

GOTV call recipients were about 14% to 26% more likely to have voted in 1998.

### **Part B**

```{r}
x_voted1996 <- c(sum(gotv_1998$voted1996), sum(non_gotv$voted1996))

n_voted1996 <- c(nrow(gotv_1998), nrow(non_gotv))

prop_test_voted1996 <- prop.test(x_voted1996, n_voted1996,conf.level = 0.95)

print(prop_test_voted1996)



```

The test shows that 71.3% of people who got GOTV calls, voted in the 1996 election vs 53% for non-call people. 

The interval does not include 0. This all indicates that people who voted in 1996 were much more likely to receive a GOTV call, and as past voting is a strong predictor of future voting. 1996 voting confounds the comparison of voting rates between the two groups. 

```{r}
x_major <- c(sum(gotv_1998$MAJORPTY), sum(non_gotv$MAJORPTY))

n_major <- c(nrow(gotv_1998), nrow(non_gotv))

prop_test_major <- prop.test(x_major, n_major,conf.level = 0.95)

print(prop_test_major)

```
The test shows that 80% of GOTV call recipients are apart of the major party compared to 74.5% of non-recipients.

The interval also does not include 0 and this all suggests that GOTV recipients are more likely to be major party members which also confounds voting rates.

```{r}
ggplot(turnout) + geom_col(aes(x=AGE,y=voted1998),fill='violet') + theme_minimal()
#cor(turnout$AGE,turnout$voted1998)

ggplot(turnout) + geom_col(aes(x=AGE,y=GOTV_call),fill='violet') + theme_minimal()
#cor(turnout$AGE,turnout$GOTV_call)


```

From the graph and correlation value of 0.30 you can see that older people are more likely to vote. Also older people were more likely to get a GOTV call. Since age is associated with both receiving a GOTV call and voting in 1998 it is confounding the voting rates. 

### **Part C**


```{r}
turn_match <- matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY, 
                     data = turnout, 
                     method = "nearest", 
                     ratio = 5)

summary(turn_match, standardize = TRUE)

turn_matched = match.data(turn_match)

summary(turn_matched)
```

After matching, the mean differences were all reduced to values below 0.1, indicating these covariates are well balanced. The summary statistics confirm that there are no significant differences in the distributions of these variables between the groups.

```{r}
gotv_matched <- turn_matched[turn_matched$GOTV_call == 1,]

non_gotv_matched <- turn_matched[turn_matched$GOTV_call == 0,]

p_treated <- mean(gotv_matched$voted1998)
p_control <- mean(non_gotv_matched$voted1998)

```

```{r}
n_match_total <- nrow(gotv_matched)
n_non_match_total <- nrow(non_gotv_matched)

diff_prop_match <- p_treated - p_control
match_diff <- sqrt((p_treated * (1 - p_treated) / n_match_total) + (p_control * (1 - p_control) / n_non_match_total))

match_lower <- diff_prop_match - z_star * match_diff
match_upper <- diff_prop_match + z_star * match_diff

```

The proportion of those receiving a GOTV call who voted in 1998 is 0.65
The sample proportion of those not receiving a GOTV call who voted in 1998 is 0.57
GOTV call recipients were about 1.3% to 14% more likely to have voted in 1998.

Overall according to these stats and the confidence interval it suggests that the GOTV calls did increase voter turnout and since this is the result after matching we can confirm it is not due to other confounding factors.
