---
title: SDS 315 - HW8"
author: "Jason George (jg82397)"
date: "https://github.com/JasonGeorge978/SDS-315"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=5, fig.width=10, warning=FALSE, message=FALSE, tidy=TRUE, echo=FALSE, tidy.opts=list(width.cutoff=60))
```

```{r}
library(ggplot2)
library(tinytex)
library(mosaic)
creatinine = read.csv('creatinine.csv')
```

\newpage

## **Problem 1: regression warm up**

```{r}
lmcreat = lm(creatclear ~ age, data=creatinine)
```


### **Part A**

```{r}
expected_rate_for_55 = 147.8129158 + (-0.6198159*55)
```
The creatinine clearance rate should we expect for a 55-year-old is 113.7. The rate should be between 113-114 and this was determined by preforming a linear regression. Then these are the calculations: 147.8129158 + (-0.6198159*55)

### **Part B**

```{r}
coef(lmcreat)
```

Creatinine clearance rate decreases by a value of -0.62 ml/minute per year with age. I got this value from the linear regression we did on the rate and age to find that value as the slope. 

### **Part C**

The 40-year-old with a rate of 135 is healthier. I determined this by doing the actual - expected of both and found the 40-year old had a increase of 11.9 from the expected. The equation for the 40 year old is `135 - (147.8129158 + (-0.6198159*40))`. The equation for the 60 year old is `112 - (147.8129158 + (-0.6198159*60))`. 

\newpage

## **Problem 2: Modeling disease growth**

```{r}
covid = read.csv('covid.csv')

italy_covid = covid[covid$country=='Italy',]

spain_covid = covid[covid$country=='Spain',]
```

```{r}
boot_italy= do(10000)*lm(log(deaths) ~ days_since_first_death,data=resample(italy_covid))

interval_italy = confint(boot_italy,level=0.95)


```


```{r}
gr_italy_lower <- round(interval_italy$lower[interval_italy$name == "days_since_first_death"],3)
gr_italy_upper <- round(interval_italy$upper[interval_italy$name == "days_since_first_death"],3)



dt_italy_lower = round(70/(gr_italy_lower*100),1)
dt_italy_upper = round(70/(gr_italy_upper*100),1)
```

### **Part 1**

The 95% confidence interval of growth rate for Italy is from 0.159 to 0.208. The 95% confidence interval of doubling time in Italy is from 3.4 to 4.4. 

```{r}
boot_spain= do(10000)*lm(log(deaths) ~ days_since_first_death,data=resample(spain_covid))

interval_spain = confint(boot_spain,level=0.95)
```


```{r}
gr_spain_lower <- round(interval_spain$lower[interval_spain$name == "days_since_first_death"],3)
gr_spain_upper <- round(interval_spain$upper[interval_spain$name == "days_since_first_death"],3)



dt_spain_lower = round(70/(gr_spain_lower*100),1)
dt_spain_upper = round(70/(gr_spain_upper*100),1)
```

The 95% confidence interval of growth rate for Spain is from 0.235 to 0.316. The 95% confidence interval of doubling time in Spain is from 2.2 to 3. 

```{r}
ggplot(covid) + geom_line(aes(x=days_since_first_death,y=deaths,color=country)) + theme_minimal()
```


\newpage

## **Problem 3: price elasticity of demand**
```{r}
milk = read.csv('milk.csv')
```

```{r}
boot_milk <- do(10000) * lm(log(sales) ~ log(price), data = resample(milk))

interval_milk = confint(boot_milk,level=0.95)

interval_milk

#interval_milk$lower[interval_milk$name == "log.price."]

#interval_milk$upper[interval_milk$name == "log.price."]


```

The 95% confidence interval for the price elasticity of the demand for milk is from -1.77 to -1.46. To estimate the elasticity I used the power law which means we can use linear regression on the log of both values. Then the bootstrap of the slope gives the elasticity. 
